---
title: 'Stat 480 - Homework #7'
author: "Carly Scott"
date: "3/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Star Wars

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.

2. We are back to a survey collected by FiveThirtyEight. This time we are interested in a survey on Star Wars - the accompanying article is published [here](https://fivethirtyeight.com/features/americas-favorite-star-wars-movies-and-least-favorite-characters/)

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
starwars <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/star-wars-survey/StarWars.csv")

# the following lines are necessary to make the second line of the file 
# part of the names:
line1 <- names(starwars)
line2 <- unlist(starwars[1,])
varnames <- paste(line1, line2)
# clean up some of the multibyte characters:
names(starwars) <- stringi::stri_trans_general(varnames, "latin-ascii")

starwars <- starwars[-1,]

names(starwars) <- enc2native(stringi::stri_trans_general(varnames, "latin-ascii"))

View(starwars)
str(starwars)
```

3. How many people respondent to the survey? How many people have seen at least one of the movies? Use the variable `Have you seen any of the 6 films in the Star Wars franchise? Response` to answer this question. Only consider responses of participants who have seen at least one of the Star Wars films for the remainder of the homework.

    1186 people responded to this survery.  
    936 people have seen at least oe of the 6 films in the Star Wars franchise.
  
```{r}
names(starwars)
summary(starwars)
library(dplyr)
library(ggplot2)

starwars %>% ggplot(aes(x = `Have you seen any of the 6 films in the Star Wars franchise? Response`)) + geom_bar()


starwars2<-starwars %>% filter(`Have you seen any of the 6 films in the Star Wars franchise? Response`=="Yes")


starwars2 %>% ggplot(aes(x = `Have you seen any of the 6 films in the Star Wars franchise? Response`)) + geom_bar()

dim(starwars2)

```

4. Variables `Gender Response` and `Age Response` are two of the demographic variables collected. Use `dplyr` to provide a frequency break down for each variable. Does the result surprise you? Comment. Reorder the levels in the variable `Age Response` from youngest to oldest.

    The gender break down for respondants to this survery are about even. There were 423 males, and 397 females, with 116 non responses. The age break down for the respondants to this survery are also pretty similar. From least frequent to most frequent age, the distribution is as follows, 18-29, >60, 30-44, 45-60. 
    The result of the variable distribution of gender is not surprising. Star Wars culture is popular with both men and women. The distribtion of age is also not surprising, the most popular age group is 45-60. The people in this age bracket were young and growing up when the first film was released in 1977, so it would make sense that that would be the most frequent age bracket.


      
```{r}
library(dplyr)


starwars2 %>% group_by(`Gender Response`) %>% tally()
starwars2 %>% group_by(`Age Response`) %>% tally()

starwars2 %>% ggplot(aes(x = `Gender Response`)) + geom_bar()
starwars2 %>% ggplot(aes(x = factor(`Age Response`, levels=c("18-29", "30-44", "45-60","> 60", "NA")))) + geom_bar()




```


5. Variables 10 through 15 answer the question: "Please rank the Star Wars films in order of preference with 1 being your favorite film in the franchise and 6 being your least favorite film." for each of the films. Use `dplyr` to find the average rank for each of the films (Note: you will first have to cast the responses to numbers). Are average ranks different between the mens' and womens' rankings?
On how many responses are the averages based? Show these numbers together with the averages.

    The ranks of the movies between genders is about the same. Men and women rated the movies about the same. 397 females and 423 males responded. Episode V: The Empire Strikes back is rated best out of the movies (rightfully so), and Episode III: Revenge of the Sith appears to be the least favorite. However, these numbers might be affected by missing values.
   

```{r}
names(starwars2)


str(starwars2)
average.gender<- starwars2 %>% 
  group_by(`Gender Response`) %>%
  summarise("Episode I" = mean(as.numeric(`Please rank the Star Wars films in order of preference with 1 being your favorite film in the franchise and 6 being your least favorite film. Star Wars: Episode I  The Phantom Menace`), na.rm= TRUE), n=n(), "Episode II" = mean(as.numeric(`X11 Star Wars: Episode II  Attack of the Clones`), na.roim = TRUE), n=n(), "Episode III" = mean(as.numeric(`X12 Star Wars: Episode III  Revenge of the Sith`), na.rm = TRUE), n=n(), "Episode IV" = mean(as.numeric(`X13 Star Wars: Episode IV  A New Hope`), na.rm = TRUE), n=n(), "Episode V" = mean(as.numeric(`X14 Star Wars: Episode V The Empire Strikes Back`), na.rm = TRUE), n=n(), "Episode VI" = mean(as.numeric(`X15 Star Wars: Episode VI Return of the Jedi`), na.rm = TRUE), n=n())


```

6. R2 D2 or C-3P0? Which of these two characters is the more popular one? Use responses to variables 25 and 26 to answer this question. Note: first you need to define  "popularity" based on the available data.  

      Popularity for this instance will be defined as the number of people who selected "very favorably" for each of the Star Wars characters. The final tally is reveals that R2 D2 is a more popular character than C-3P0. However, the number of people who find R2 D2 favorable is very close to the number of people who find C-3P0 favorable. R2 D2 beat C-3P0 in the popularity contest, 562 to 474. Rightfully so, R2D2 is a loveable robot and C-3P0 is a neurotic droid. 

```{r}

starwars2 %>% group_by(`X26 R2 D2`) %>% tally()
starwars2 %>% group_by(`X25 C-3P0`) %>% tally() 

starwars2 %>% group_by(`X26 R2 D2`) %>% summarise(n = length(`RespondentID NA`))
starwars2 %>% group_by(`X25 C-3P0`) %>% summarise(n = length(`RespondentID NA`))

starwars2 %>% ggplot(aes(x = `X26 R2 D2`)) + geom_bar()
starwars2 %>% ggplot(aes(x = `X25 C-3P0`)) + geom_bar()

starwars2 %>% group_by(`X27 Jar Jar Binks`)%>% summarise(n = length(`RespondentID NA`))
starwars2 %>% ggplot(aes(x = `Please state whether you view the following characters favorably, unfavorably, or are unfamiliar with him/her. Han Solo`)) + geom_bar()

```

Due date: the homework is due before class on Thursday. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html file with it. 
